Elephant Seal Age JAE Intraspecific Variation

### 2023-11-15 paper/paperv2.qmd
Streamlining the html site

Panel B shown with axis constrained below.

```{r}
#| label: fig-age_dist2
#| fig-cap: "B: Percentage of each age class that were observed breeding, with axes constrained."
#| fig-width: 9

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(n_observed > 2) %>% 
  ggplot(aes(x = age, y = perc_breed)) + 
  stat_smooth(method = "lm", 
              formula = y ~ poly(x , 2), 
              se = TRUE, 
              color = "#849ED5", 
              fill = "#C8D3ED") +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(.25, .75)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
  theme_few()

breed_nonbreed  
```

@@

**Note:** *Try to plot these quadraticlly??*

@@

**Note**: *Getting error "boundary (singular) fit" - possibly not enough data to fit this model? I.e., too few recruits. Also, need to apply 7 year filter to this one too????*

@@

@@

If not maternal effect senescence, then perhaps a strategy shift?

@@

We found that as maternal age increases, pup sex skews slightly towards males.

@@

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 20) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_hline(yintercept = 0.5, lty = 2, color = "#829BD1") +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

@@

```{r}
pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
                     data = filter(sealdat, 
                                   pupsex %in% c("M", "F"), 
                                   age < 18), 
                     family = "binomial")
summary(pup_sex_mod)
```

@@

**Note:** *Add in illustrator the colors between each phenological variable, so people can understand the timing. Fix the labels on panel B and write a caption.*

```{r}
#| label: fig-phenology
#| fig-height: 7
#| fig-cap: "Phenology caption"

>>>>

@@

@@

@@

We also plotted survival bias to show that 95% of pups were resighted within 7 years. Both plots are supplemental.

@@

@@

We found that as maternal age increases, pup sex skews slightly towards males - but it's not significant.

@@

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 19) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_hline(yintercept = 0.5, lty = 2, color = "#829BD1") +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

@@

```{r}
pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
                     data = filter(sealdat, 
                                   pupsex %in% c("M", "F"), 
                                   age < 19), 
                     family = "binomial")
summary(pup_sex_mod)
```

@@

The table is supplemental.

```{r}
#| label: fig-phenology
#| fig-height: 7
#| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles. A: Visualization of life history timing across ages, faceted by breeding status. White dashed lines on each panel represent the raw data from the other panel to facilitate comparisons between breeders and non-breeders. B: LMM coefficients for phenological parameters. The units for maternal age coefficients are in days per year. The coefficients for breeding status are in days. BA = Breeding arrival; BD = Breeding departure; MA = Molt arrival; MD = Molt departure. Black points are significant, gray points are not significant."
